---
title: "RMarkdown Wordpress Template"
author: "3.14a"
date: "Monday, March 09, 2015"
output: html_document
---

This is R Markdown Wordpress template file.

Run *wordpress* chunk with option `eval=FALSE,echo=FALSE` to publish this R Markdown file on your Wordpress website. Eval is set to FALSE, otherwise the posting will be trapped in a loop.

```{r wordpress,eval=FALSE,echo=FALSE}

# replace <password> <url> make sure chunk option is set to echo=FALSE !
if (!require('RWordPress')){install.packages('RWordPress', repos = 'http://www.omegahat.org/R', type = 'source')}


library(RWordPress)

options(WordpressLogin = c(nurakawa = 'p0mpeiib@still3'), WordpressURL = 'https://nurakawa.wordpress.com/xmlrpc.php')

library(knitr)

# Knitr options: upload plots/images to wordpress

opts_knit$set(upload.fun = function(file){library(RWordPress);uploadFile(file)$url;})

# enable toc (comment out if not needed)
library(markdown)
options(markdown.HTML.options =  c(markdownHTMLOptions(default = T),"toc"))

# Upload featured image / post thumbnail: option: wp_post_thumbnail=postThumbnail$id
postThumbnail <- RWordPress::uploadFile("figure/post_thumbnail.png",overwrite = TRUE)

postid <- knit2wp('sample-wordpress.Rmd', action = c("newPost"),title = 'RMarkdown Wordpress Template',categories=c('R'),mt_keywords = c('R','RMarkdown'),publish=FALSE) # add featured image include: wp_post_thumbnail=postThumbnail$id 
```

## Plot

The following plot shows the annual blog frequencies and serves as example of a succesful file upload.

```{r, fig.width=12,fig.height=4,fig.cap="Blog posts by year", eval=FALSE}
library(ggplot2)
ggplot(mtcars, aes(factor(cyl)))+geom_bar()
```

```{r, eval=TRUE}
plot(iris$Sepal.Length, iris$Sepal.Width)
```



## Math

Test the Math equation rendering:

Inline formula with $a=3$ and $b=5$

$$ x = a + b $$


## Table


```{r, echo=FALSE, results='asis'}
knitr::kable(data.frame(a=1:5,b=rnorm(5)))
```